---
title: "College Tuition, Diversity, and Pay"
author: "Joshua Cook"
date: "3/10/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidymodels)
library(magrittr)
library(tidyverse)

theme_set(theme_minimal())

set.seed(0)
```

## Data preparation

```bash
cd data
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv'
cd ..
```

Today, I would specifically like to employ a linear modex effects model using the `lmer()` function from the 'lme4' package.
The question I ask and data set I use will be targeted towards being able to implement these tools.

I think I will use the "tuition_income.csv" data:

|variable    |class     |description |
|:-----------|:---------|:-----------|
|name        |character | School name |
|state       |character | State Name |
|total_price |double    | Total price in USD |
|year        |double    | year |
|campus      |character | On or off-campus |
|net_cost    |double    | Net-cost - average actually paid after scholarship/award |
|income_lvl  |character | Income bracket |

```{r}
tuition_income <- read_csv(file.path("data", "tuition_income.csv")) %>%
    janitor::clean_names()
```

```{r}
tuition_income %>%
    skimr::skim()
```

The `income_lvl` data is currently a categorical variable, but I would like to model the income as a function of the rest of the variables.
Therefore, I will make it a continuous variable by randomly sampling from a uniform distribution within the range.

```{r}

income_range <- tibble::tribble(
    ~income_lvl, ~income_min, ~income_max,
    "0 to 30,000", 0, 30e3,
    "30,001 to 48,000", 30e3, 48e3,
    "48_001 to 75,000", 48e3, 75e3,
    "75,001 to 110,000", 75e3, 110e3,
    "Over 110,000", 110e3, 150e3
)

tuition_income %<>%
    left_join(income_range, by = "income_lvl") %>%
    group_by(income_lvl) %>%
    mutate(
        n = n(),
        income = runif(n = unique(n), 
                       min = unique(income_min),
                       max = unique(income_max)),
        income = round(income)) %>%
    ungroup()
```

To limit the scope of the analysis (and actually fit models in a reasonable amount of time), I will limit the data to just the years 2015 - 2018 and the schools to those in California and Massacheusettes.

```{r}
tuition_income %<>%
    filter(state %in% c("MA", "CA")) %>%
    filter(between(year, 2015, 2018))
```

I'm going to further restrict the analysis to a handful of schools, again, just ot limit the score of the analysis.

```{r}
schools <- c(
    "Stanford",
    "California Institute of Technology",
    "Harvard University",
    "Massachusetts Institute of Technology",
    "Boston University",
    "Boston College",
    "Tufts University"
)

tuition_income %<>%
    filter(
      str_detect(name, "University of California-") | name %in% schools
    ) %>%
    mutate(
      name = str_replace(name, "University of California-", "UC "),
      name = str_replace(name, "Massachusetts Institute of Technology", "MIT"),
      name = str_replace(name, "California Institute of Technology", "CalTech")
    )
```


## EDA

```{r}
tuition_income %>%
    ggplot(aes(x = total_price, y = net_cost)) +
    geom_density_2d(color = "black", alpha = 0.6, size = 0.6) +
    geom_point(aes(color = campus), size = 0.6, alpha = 0.3) +
    scale_x_continuous(limits = c(0, 75000), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 75000), expand = c(0, 0)) +
    scale_color_brewer(palette = "Set1")
```

```{r}
tuition_income %>%
    ggplot(aes(x = income)) +
    geom_density(aes(color = income_lvl, fill = income_lvl)) +
    scale_color_brewer() +
    scale_fill_brewer() +
    labs(color = "income level", fill = "income level")
```

```{r}
tuition_income %>%
  ggplot(aes(x = net_cost, y = income)) +
  geom_point(aes(color = name, shape = state)) +
  geom_smooth() +
  scale_color_manual(
    values = randomcoloR::distinctColorPalette(n_distinct(tuition_income$name)),
    guide = guide_legend(ncol = 2, order = 0)
  )
```


## Model

```{r}
income_data <- tuition_income %>%
    ungroup() %>%
    select(income, name, state, year, campus, net_cost) %>%
    mutate(
        income = scale(income)[, 1],
        net_cost = scale(net_cost)[, 1],
        year = scale(year)[, 1]
    )

head(income_data)
```

```{r}
dim(income_data)
```

### Linear model

```{r}
income_fit1 <- lm(income ~ ., data = income_data)
summary(income_fit1)
```

```{r}
plot(income_fit1)
```

### Generalized linear model

```{r}
income_fit2 <- glm(income ~ ., data = income_data)
summary(income_fit2)
```

```{r}
plot(income_fit2)
```

```{r}
anova(income_fit1, income_fit2)
```


### Linear mixed effects model

```{r}
income_fit3 <- lmer(income ~ year + net_cost + campus + (1 | state/name), 
                    data = income_data)
summary(income_fit3)
```

```{r}
plot(income_fit3, type = c("p", "smooth"))
```

```{r}
lattice::qqmath(income_fit3, id = 0.05,)
```

```{r}
confint(income_fit3)
```

