---
title: "Best Rap Artists"
author: "Joshua Cook"
date: "4/14/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(conflicted)
library(glue)
library(magrittr)
library(tidygraph)
library(ggraph)
library(tidyverse)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

theme_set(theme_minimal())
```

## Data

```bash
cd data
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv'
cd ..
```

```{r}
rap_polls <- read_csv(file.path("data", "polls.csv")) %>%
    janitor::clean_names() %>%
    select(-critic_country2)
rap_rankings <- read_csv(file.path("data", "rankings.csv")) %>%
    janitor::clean_names()
```

```{r}
rap_rankings %>% 
    count(title, artist) %>% 
    filter(n > 1)
```


The song *All Of The Lights* by Kanye West was split into two songs for some reason in the polls, but I manually combined the data and reevaluated the rankings.

```{r}
kw_all_of_the_light <- rap_rankings %>% 
    filter(title == "All Of The Lights") %>%
    mutate(gender = "mixed") %>%
    group_by(title, artist, year, gender) %>%
    summarise(points = sum(points),
              n = sum(n),
              n1 = sum(n1),
              n2 = sum(n2),
              n3 = sum(n3),
              n4 = sum(n4),
              n5 = sum(n5)) %>%
    ungroup() %>%
    mutate(id = 0) %>%
    select(all_of(colnames(rap_rankings))) %>%
    as_tibble()

rap_rankings %<>%
    filter(title != "All Of The Lights") %>%
    bind_rows(kw_all_of_the_light)
```


## EDA

### `rap_polls`

```{r}
rap_polls %>%
    count(gender) %>%
    ggplot(aes(x = gender, y = n)) +
    geom_col(aes(fill = gender)) +
    theme(legend.position = "none",
          axis.title.x = element_blank()) +
    labs(y = "count",
         title = "Distribution of genders in top-rated Rap songs")
```

```{r}
rap_polls %>%
    count(year) %>%
    ggplot(aes(x = year, y = n)) +
    geom_point() +
    geom_line(alpha = 0.5) +
    geom_smooth(color = "black") +
    scale_x_continuous(expand = c(0, 0)) +
    labs(x = "year",
         y = "number of songs included in critic rankings",
         title = "The number of songs critics included in their rankings per year")
```

```{r}
rap_polls %>%
    count(title, artist, sort = T) %>%
    top_n(20, wt = n) %>%
    mutate(lbl = paste0(artist, ": ", title),
           lbl = fct_reorder(lbl, n)) %>%
    ggplot(aes(x = n, y = lbl)) +
    geom_col() +
    scale_x_continuous(expand = expansion(mult = c(0, 0.02))) +
    theme(axis.title.y = element_blank()) +
    labs(x = "count",
         title = "Songs most frequently included in\ncritic top-5 rankings")
```

### `rap_rankings`

```{r}
rap_rankings %>%
    select(title, artist, year, points) %>%
    set_names(c("Song", "Artist", "Year", "Points")) %>%
    head(n = 10) %>%
    knitr::kable()
```

```{r}
rap_rankings %>%
    distinct(title, artist) %>%
    count(artist) %>%
    top_n(20, wt = n) %>%
    mutate(artist = fct_reorder(artist, n)) %>%
    ggplot(aes(x = n, y = artist)) +
    geom_col() +
    scale_x_continuous(expand = expansion(mult = c(0, 0.02))) +
    theme(axis.title.y = element_blank()) +
    labs(x = "count",
         title = "Number of songs per artist in rankings")
```

## Graph

```{r}
rap_rankings_select <- rap_rankings %>%
    select()
rap_polls_select <- rap_polls %>%
    select(-gender, -year) %>%
    rename(critic_rank = rank)
```

```{r}
rap_data <- rap_rankings %>%
    select(-(n1:n5)) %>%
    left_join(rap_polls_select, by = c("title", "artist")) %>%
    mutate(title = ifelse(
        title == "Hip Hop", glue("{title} ({artist})"), title
    ))

rap_data
```

```{r}
rap_el_1 <- rap_data %>%
    select(critic_name, title) %>%
    set_names(c("from", "to"))
rap_el_2 <- rap_data %>%
    select(artist, title) %>%
    set_names(c("from", "to"))


rap_song_data <- rap_data %>%
    distinct(title, year, points)

rap_artist_data <- rap_data %>%
    distinct(title, artist, year, gender, points) %>%
    group_by(artist) %>%
    summarise(total_points = sum(points),
              num_songs = n_distinct(title)) %>%
    ungroup()

rap_critic_data <- rap_data %>%
    distinct(critic_name, critic_rols, critic_rols)

rap_graph <- as_tbl_graph(bind_rows(rap_el_1, rap_el_2)) %N>%
    mutate(type = case_when(
        name %in% rap_data$title ~ "song_title",
        name %in% rap_data$artist ~ "song_artist",
        name %in% rap_data$critic_name ~ "critic_name"
    )) %N>%
    left_join(rap_song_data, by = c("name" = "title")) %>%
    left_join(rap_artist_data, by = c("name" = "artist")) %>%
    left_join(rap_critic_data, by = c("name" = "critic_name"))

rap_graph
```

Three components in total, and two are far smaller than the rest.

```{r}
# Number of components.
igraph::count_components(rap_graph)

# Size of components.
igraph::decompose(rap_graph) %>% map_dbl(igraph::vcount)
```

These subcomponents represent two different critics with completely unique lists of favorite songs.
Further, their songs are all from artists without any other ranked songs.

```{r}
rap_graph %N>%
    jhcutils::filter_component_size(min_size = 0, max_size = 15) %>%
    ggraph(layout = "stress") +
    geom_edge_link(alpha = 0.4, color = "grey50", width = 1.3) +
    geom_node_point(aes(color = type), size = 3) +
    geom_node_text(aes(label = name, color = type), 
                   repel = TRUE, family = "Arial") +
    scale_color_brewer(palette = "Set1") +
    theme_graph()
```

```{r, fig.height=8, fig.width=8}
rap_graph %N>%
    jhcutils::get_giant_component() %>%
    mutate(centrality = centrality_betweenness(directed = FALSE),
           lbl = ifelse(centrality > quantile(centrality, 0.95), name, NA)) %E>%
    mutate(centrality = centrality_edge_betweenness(directed = FALSE)) %N>%
    ggraph(layout = "stress") +
    geom_edge_link(aes(alpha = centrality), color = "grey50", width = 1) +
    geom_node_point(aes(color = type, size = centrality), alpha = 0.7) +
    geom_node_text(aes(label = lbl), color = "grey25",
                   repel = TRUE, size = 3, family = "Arial") +
    scale_color_brewer(palette = "Set1") +
    scale_edge_alpha_continuous(range = c(0.2, 0.9), guide = NULL) +
    scale_size_continuous(range = c(0.8, 6), guide = NULL) +
    scale_alpha_continuous(range =c(0.2, 0.8), guide = NULL) +
    theme_graph() +
    theme(legend.position = "bottom")
```

Potentially better visualization by reducing the graph down to the minimum spanning tree.

```{r, fig.height=8, fig.width=8}
rap_graph %N>%
    jhcutils::get_giant_component() %>%
    mutate(centrality = centrality_betweenness(directed = FALSE),
           lbl = ifelse(centrality > quantile(centrality, 0.95), name, NA)) %E>%
    mutate(centrality = centrality_edge_betweenness(directed = FALSE)) %N>%
    convert(to_minimum_spanning_tree) %>%
    ggraph(layout = "stress") +
    geom_edge_link(aes(alpha = centrality), color = "grey50", width = 1) +
    geom_node_point(aes(color = type, size = centrality), alpha = 0.7) +
    geom_node_text(aes(label = lbl), color = "grey25",
                   repel = TRUE, size = 3, family = "Arial") +
    scale_color_brewer(palette = "Set1") +
    scale_edge_alpha_continuous(range = c(0.2, 0.9), guide = NULL) +
    scale_size_continuous(range = c(0.8, 6), guide = NULL) +
    scale_alpha_continuous(range =c(0.2, 0.8), guide = NULL) +
    theme_graph() +
    theme(legend.position = "bottom")
```

Group the critics by clustering the graph.

```{r, fig.height=8, fig.width=8}
clustered_rap_graph <- rap_graph %N>%
    jhcutils::get_giant_component() %>%
    convert(to_undirected) %>%
    mutate(cls = group_spinglass()) %E>%
    filter(.N()$cls[from] == .N()$cls[to])

clustered_rap_graph %>%
    ggraph(layout = "stress") +
    geom_edge_link(alpha = 0.4, color = "grey50", width = 1) +
    geom_node_point(aes(color = type), size = 1.3, alpha = 0.7) +
    scale_color_brewer(palette = "Set1") +
    theme_graph() +
    theme(legend.position = "bottom")
```

The critics clustered based on their relationships in the graph.

```{r}
clustered_rap_graph %N>%
    as_tibble() %>%
    select(name, type, cls) %>%
    filter(type == "critic_name") %>%
    arrange(cls) %>%
    group_by(cls) %>%
    summarise(critics = paste0(name, collapse = ", ")) %>%
    knitr::kable()
```


### Future directions

Here are some of the ways this type of data structure could be useful:

1. Using the closeness on the graph to recommend songs or artists to the people based on songs they already know and like.
2. Finding the core rap songs and artists.
3. Grouping the critics, artists, or songs by using a graph cutting or clustering algorithm.
4. Predicting similarity of artists, songs, and critics.
