---
title: "Beer Production"
author: "Joshua Cook"
date: "3/31/2020"
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

Download the data from the GitHub [page](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-31/readme.md)

```bash
cd data

wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv'

cd ..
```

Here is the brief summary we are given about the data:

>The data this week comes from the Alcohol and Tobacco Tax and Trade Bureau (TTB). H/t to Bart Watson for sharing the source of the data.
>
> There's a literal treasure trove of data here:
>
>  - State-level beer production by year (2008-2019)
>  - Number of brewers by production size by year (2008-2019)
>  - Monthly beer stats aggregated across the US (2008-2019)
>  
> Some considerations:
> 
>  - A barrel of beer for this data is 31 gallons
>  - Most data is in barrels removed/taxed or produced
>  - Removals = "Total barrels removed subject to tax by the breweries comprising the named strata of data", essentially how much was produced and removed for consumption.
>  - A LOT of data came from PDFs - I included all the code I used to grab data and tidy it up, take a peek and try out your own mechanism for getting the tables out.

I have seen a few people looking at the "brewing_materials.csv" data, so I think I will instead analyze the "brewer_size.csv" today.

```{r}
library(magrittr)
library(tidyverse)

theme_set(theme_minimal())
```


```{r}
brewer_size <- read_csv(file.path("data", "brewer_size.csv"))
brewer_size
```

There are some categories of brewer size that overlap.
I manually curated the list to have only non-overlapping groups.

```{r}
sort(unique(brewer_size$brewer_size))
```

```{r}
brewer_size %<>%
    mutate(brewer_size = case_when(
        brewer_size == "Total" ~ NA_character_,
        brewer_size == "1,000,000 to 6,000,000 Barrels" ~ "1,000,001 to 6,000,000 Barrels",
        brewer_size == "1,000,001 to 1,999,999 Barrels" ~ "1,000,001 to 6,000,000 Barrels",
        brewer_size == "2,000,000 to 6,000,000 Barrels" ~ "1,000,001 to 6,000,000 Barrels",
        TRUE ~ brewer_size
    ))
```

I then declared the order of the `brewer_size` column.

```{r}
brewer_size_order <- c(
    "Zero Barrels",
    "Under 1 Barrel",
    "1 to 1,000 Barrels",
    "1,001 to 7,500 Barrels",
    "7,501 to 15,000 Barrels",
    "15,001 to 30,000 Barrels",
    "30,001 to 60,000 Barrels",
    "60,001 to 100,000 Barrels",
    "100,001 to 500,000 Barrels",
    "500,001 to 1,000,000 Barrels",
    "1,000,001 to 6,000,000 Barrels",
    "6,000,001 Barrels and Over"
)

brewer_size %<>%
    mutate(brewer_size = factor(brewer_size, levels = brewer_size_order)) %>%
    group_by(year, brewer_size) %>%
    summarise(n_of_brewers = sum(n_of_brewers)) %>%
    ungroup()
```


## EDA

The first plot shows the total number of companies at each size.

```{r}
brewer_size %>%
    filter(!is.na(brewer_size)) %>%
    ggplot(aes(x = year, y = n_of_brewers, group = brewer_size, color = brewer_size)) +
    geom_line() +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(x = "year",
         y = "number of brewers",
         color = "brewer size")
```

The plot below is the same as above, but after removing the top two largest groups.

```{r}
brewer_size %>%
    filter(!is.na(brewer_size)) %>%
    filter(!brewer_size %in% c("1 to 1,000 Barrels", "1,001 to 7,500 Barrels")) %>%
    ggplot(aes(x = year, y = n_of_brewers, group = brewer_size, color = brewer_size)) +
    geom_line() +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(x = "year",
         y = "number of brewers",
         color = "brewer size")
```

The trend in the "Under 1 Barrel" group is rather interesting.
They grow sharply and then dip in 2017.
The following analysis will try to tell if this is because they are going bankrupt, or are expanding into higher groups.


## The "Under 1 Barrel" drop

We can start by analyzing the year-over-year change in number of companies, i.e. the dirst derivative of the lines above.

```{r}
calc_year_over_year <- function(x) {
    y <- x
    for (i in seq(1, length(x))) {
        if (i == 1) {
            y[[i]] <- 0
        } else {
            y[[i]] <- x[[i]] - x[[i - 1]]
        }
    }
    return(y)
}


brewer_size %<>%
    filter(!is.na(brewer_size)) %>%
    group_by(brewer_size) %>%
    arrange(year) %>%
    mutate(yoy_change = calc_year_over_year(n_of_brewers)) %>%
    ungroup()
```

We can see that a large drop in the "Under 1 Barrel" category is concomittant with an increase in the "1 to 1,000 Barrels" group. 
The difference does account for the change: from 2017 to 2018, the "Under 1 Barrel" category lost 400 companies, but the "1 to 1,000 Barrels" category gained 600.
So not only does it look like many small operations scaled up, but additional new companies were started and went instantly to producing more than one barrel.
The same trend holds for the 2018 to 2019 transition.

```{r}
brewer_size %>%
    ggplot(aes(x = year, y = yoy_change, group = brewer_size, color = brewer_size)) +
    geom_line() +
    scale_x_continuous(breaks = scales::pretty_breaks()) +
    labs(x = "year",
         y = "year-over-year change",
         color = "brewer size")
```

```{r}
brewer_size %>%
    filter(year >= 2016) %>%
    filter(brewer_size %in% c("Under 1 Barrel", "1 to 1,000 Barrels")) %>%
    mutate(year = factor(year)) %>%
    ggplot(aes(x = year, y = yoy_change, group = brewer_size, color = brewer_size)) +
    geom_hline(yintercept = 0, color = "grey50", size = 0.6) +
    geom_line() +
    labs(x = "year",
         y = "year-over-year change",
         color = "brewer size")
```

```{r}
brewer_size %>%
    filter(year >= 2016) %>%
    filter(brewer_size %in% c("Under 1 Barrel", "1 to 1,000 Barrels")) %>%
    group_by(year) %>%
    arrange(brewer_size) %>%
    summarise(diff_yoy_change = sum(yoy_change)) %>%
    ungroup() %>%
    ggplot(aes(x = year, y = diff_yoy_change)) +
    geom_col() +
    labs(x = "year",
         y = "sum of year-over-year")
```

