---
title: "Traumatic Brain Injury"
author: "Joshua Cook"
date: "3/24/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

Download the data from the GitHub [page](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-31/readme.md)

```bash
cd data

wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_year.csv'
wget 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_military.csv'

cd ..
```

I think we will use the `tbi_age.csv` data.
It provides the estimated number and rate (per 100,000) of injuries for various age gruops in 2014.

```{r}
library(factoextra)
library(patchwork)
library(tidyverse)

theme_set(theme_minimal())
```


```{r}
tbi_age <- read_csv(file.path("data", "tbi_age.csv")) %>%
    janitor::clean_names()

head(tbi_age)
```

There is a "0-17" age group and then smaller sub-groups in the same range, but I don't know if they are actual subsets of the "0-17" group or there is something weird with the data set.

```{r}
tbi_age %>%
    filter(
        type == "Emergency Department Visit" & 
        injury_mechanism == "Unintentional Falls" &
        age_group %in% c("0-4", "5-14")
    ) %>%
    pull(number_est) %>%
    sum()
```

```{r}
tbi_age %>%
    filter(
        type == "Emergency Department Visit" & 
        injury_mechanism == "Unintentional Falls" &
        age_group == "0-17"
    ) %>%
    pull(number_est)
```

The 0-17 group had 397,190 vistis to the ED due to unintentional falls and the 0-4 and 5-14 groups had 363,860 visits for the same reason, combined.
Thus, they are likely subsets of the 0-17 group.
This is important to avoid double counting.


## EDA

We can plot a PCA of the age groups in a space with axes for the injury mechanism.
Since the rate estimated is per 100,000 people in the age group, we can sum the `rate_est` values within each age group and injury mechanism.

```{r}
injury_mech <- tbi_age %>%
    filter(!age_group %in% c("Total", "0-17")) %>%
    select(age_group, injury_mechanism, rate_est) %>%
    group_by(age_group, injury_mechanism) %>%
    summarise(rate_est = sum(rate_est)) %>%
    ungroup() %>%
    pivot_wider(age_group,
                names_from = injury_mechanism,
                values_from = rate_est) %>%
    mutate_all(~ ifelse(is.na(.), 0, .)) %>%
    as.data.frame(stringsAsFactors = FALSE) %>%
    column_to_rownames("age_group")

head(injury_mech)
```

```{r}
injury_mech_pca <- prcomp(injury_mech, scale = TRUE, center = TRUE)
injury_mech_pca
```

```{r}
summary(injury_mech_pca)
```

```{r}
fviz_eig(injury_mech_pca)
```

```{r}
tibble(
    pc = c(0, seq(1, ncol(injury_mech_pca$x))),
    cum_prop = c(0, cumsum(injury_mech_pca$sdev^2 / sum(injury_mech_pca$sdev^2)))
) %>%
    mutate(label = paste0(round(cum_prop * 100), "%")) %>%
    ggplot(aes(x = pc, y = cum_prop)) +
    geom_hline(yintercept = 0, color = "grey70") +
    geom_vline(xintercept = 0, color = "grey70") +
    geom_line(linetype = 2) +
    geom_point(size = 1.2) +
    geom_text(aes(label = label), family = "Arial", 
              nudge_x = 0.2, nudge_y = -0.03) +
    labs(x = "principal component",
         y = "cumulative proportion")
```

We can see that the first three dimensions contain 97% of the information in the new injury mechanism space.

From the plots below, The first dimension is defined by "Unintentional Falls" in one direction and "Intentional self-harm", "Assault", and "Motor Vehicle Crashes" in the opposite direction, mainly separating the very old from the young adults and teenagers.
The second dimension is defined by "Unintentionally struck by or against an object" and "Other unintentional injury, mechanism unspecified" and strifies the elderly and the young (except for the 75+ and 25-34 age groups).
The third dimension is a combination of these factors and is not as opinonated at the first two.

```{r}
fviz_pca(injury_mech_pca, axes = c(1, 2))  
```

```{r}
fviz_pca(injury_mech_pca, axes = c(1, 3))
```
