---
title: "Beach Volleyball"
author: "Joshua Cook"
date: "May 19, 2020"
output: github_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")

library(glue)
library(naniar)
library(lubridate)
library(patchwork)
library(magrittr)
library(broom)
library(yardstick)
library(tidyverse)
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

theme_set(theme_minimal())
```

## Data

```{r}
vb_matches <- read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-19/vb_matches.csv",
    guess_max = 76000,
    col_types = cols()
) %>%
    janitor::clean_names()
```

## EDA

```{r}
skimr::skim(vb_matches)
```

## Prepare model data

Prepare data to use for modeling wins and losses.
Only gameplay statistics and duration will be used for the model.

```{r}
vb_match_stats <- vb_matches %>%
    select(tournament:match_num,
           w_rank, l_rank, score, duration,
           w_p1_tot_attacks:w_p1_tot_errors, w_p1_tot_aces:w_p1_tot_digs,
           w_p2_tot_attacks:w_p2_tot_errors, w_p2_tot_aces:w_p2_tot_digs,
           l_p1_tot_attacks:l_p1_tot_errors, l_p1_tot_aces:l_p1_tot_digs,
           l_p2_tot_attacks:l_p2_tot_errors, l_p2_tot_aces:l_p2_tot_digs) %>%
    filter_all(all_vars(!is.na(.))) %>%
    mutate(match_idx = row_number())
```

The gameplay statstics are summed for each team.

```{r}
vb_data <- vb_match_stats %>%
    mutate(
        w_attacks = w_p1_tot_attacks + w_p2_tot_attacks,
        w_kills = w_p1_tot_kills + w_p2_tot_kills,
        w_errors = w_p1_tot_errors + w_p2_tot_errors,
        w_aces = w_p1_tot_aces + w_p2_tot_aces,
        w_serveerrors = w_p1_tot_serve_errors + w_p2_tot_serve_errors,
        w_blocks = w_p1_tot_blocks + w_p2_tot_blocks,
        w_digs = w_p1_tot_digs + w_p2_tot_digs,
        l_attacks = l_p1_tot_attacks + l_p2_tot_attacks,
        l_kills = l_p1_tot_kills + l_p2_tot_kills,
        l_errors = l_p1_tot_errors + l_p2_tot_errors,
        l_aces = l_p1_tot_aces + l_p2_tot_aces,
        l_serveerrors = l_p1_tot_serve_errors + l_p2_tot_serve_errors,
        l_blocks = l_p1_tot_blocks + l_p2_tot_blocks,
        l_digs = l_p1_tot_digs + l_p2_tot_digs
    ) %>%
    select(match_idx, tournament:match_num,
           w_rank, l_rank, score, duration,
           w_attacks:l_digs) %>%
    mutate(l_rank = as.numeric(str_extract(l_rank, "^[:digit:]+(?=,|$)")),
           w_rank = as.numeric(str_extract(w_rank, "^[:digit:]+(?=,|$)")),
           rank_diff = w_rank - l_rank) %>%
    filter(!is.na(l_rank) & !is.na(w_rank))
```

```{r}
scale2 <- function(x, na.rm = FALSE) {
    (x - mean(x, na.rm = na.rm)) / sd(x, na.rm = na.rm)
}

d_all <- vb_data %>%
    mutate(duration = as.numeric(duration) / 60) %>%
    select(duration, w_attacks:l_digs)

d_winners <- d_all %>% 
    select(duration, w_attacks:w_digs) %>% 
    add_column(result = 1)
d_losers <- d_all %>% 
    select(duration, l_attacks:l_digs) %>% 
    add_column(result = 0)
colnames(d_winners) <- str_remove_all(colnames(d_winners), "^w_")
colnames(d_losers) <- str_remove_all(colnames(d_losers), "^l_")
d <- bind_rows(d_winners, d_losers) %>%
    mutate_at(c("duration", "attacks", "kills", "errors", "aces", 
                "serveerrors", "blocks", "digs"), scale2)

```

The modeling data was split into testing and training.

```{r}
set.seed(0)
test_idx <- sample(1:nrow(d), 0.2 * nrow(d), replace = FALSE)
test_data <- d[test_idx, ]
train_data <- d[!(1:nrow(d) %in% test_data), ]
```

It is important to check for colinearity of predictors.

```{r}
train_data %>% 
    select(-duration, -result) %>%
    sample_n(100) %>%
    pairs()
```

## Modeling Wins vs. Losses

### Model 1. Just an intercept

This model will serve as the baseline: what would be expected given the average results of games.
Since there is always one winner and one loser, the intercept should be 0.5.

```{r}
m1 <- glm(result ~ 1, data = train_data, family = "binomial")
```

### Model 2. All gameplay statistics

```{r}
m2 <- glm(result ~ attacks + kills + errors + aces + serveerrors + blocks,
          data = train_data,
          family = "binomial")
summary(m2)
```

```{r}
performance::check_collinearity(m2)
```

### Model 3. All gameplay statistics and their interaction with match duration

```{r}
m3 <- glm(result ~ duration * ., data = train_data, family = "binomial")
summary(m3)
```

```{r}
performance::check_collinearity(m3)
```

## Model evaluation

### Predictions on training and testing data

```{r}
m1_train_pred <- predict(m1, type = "response")
m1_test_pred <- predict(m1, newdata = test_data, type = "response")

m2_train_pred <- predict(m2, type = "response")
m2_test_pred <- predict(m2, newdata = test_data, type = "response")

m3_train_pred <- predict(m3, type = "response")
m3_test_pred <- predict(m3, newdata = test_data, type = "response")
```

```{r}
tibble(name = c("m1_train", "m1_test", 
                "m2_train", "m2_test", 
                "m3_train", "m3_test"),
       value = c(list(m1_train_pred),
                 list(m1_test_pred),
                 list(m2_train_pred),
                 list(m2_test_pred),
                 list(m3_train_pred),
                 list(m3_test_pred))
       ) %>%
    unnest(value) %>%
    ggplot(aes(x = value, color = name, fill = name)) +
    geom_density(alpha = 0.2, size = 1) +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")
```

### Accuracy

$$
\text{Accuracy} = \frac{TP + TN}{TP + FP + FN + TN}
$$

```{r}
my_accuracy <- function(truth, estimate) {
    accuracy(tibble(truth = factor(truth, levels = c(0, 1)),
                estimate = factor(round(estimate), levels = c(0, 1))),
         truth, estimate)
}


bind_rows(
    my_accuracy(train_data$result, m1_train_pred),
    my_accuracy(test_data$result, m1_test_pred),
    my_accuracy(train_data$result, m2_train_pred),
    my_accuracy(test_data$result, m2_test_pred),
    my_accuracy(train_data$result, m3_train_pred),
    my_accuracy(test_data$result, m3_test_pred)
) %>%
    mutate(name = c("m1 train", "m1 test", "m2 train", "m2 test", "m3 train", "m3 test"))

```

### Precision

$$
\text{Precision} = \frac{TP}{TP + FP}
$$

```{r}
my_precision <- function(truth, estimate) {
    precision(tibble(truth = factor(truth, levels = c(0, 1)),
                     estimate = factor(round(estimate), levels = c(0, 1))),
              truth, estimate)
}


bind_rows(
    my_precision(train_data$result, m1_train_pred),
    my_precision(test_data$result, m1_test_pred),
    my_precision(train_data$result, m2_train_pred),
    my_precision(test_data$result, m2_test_pred),
    my_precision(train_data$result, m3_train_pred),
    my_precision(test_data$result, m3_test_pred)
) %>%
    mutate(name = c("m1 train", "m1 test", "m2 train", "m2 test", "m3 train", "m3 test"))
```

### Recall

$$
\text{Recall} = \frac{TP}{TP + FN}
$$

```{r}
my_recall <- function(truth, estimate) {
    recall(tibble(truth = factor(truth, levels = c(0, 1)),
                  estimate = factor(round(estimate), levels = c(0, 1))),
           truth, estimate)
}


bind_rows(
    my_recall(train_data$result, m1_train_pred),
    my_recall(test_data$result, m1_test_pred),
    my_recall(train_data$result, m2_train_pred),
    my_recall(test_data$result, m2_test_pred),
    my_recall(train_data$result, m3_train_pred),
    my_recall(test_data$result, m3_test_pred)
) %>%
    mutate(name = c("m1 train", "m1 test", "m2 train", "m2 test", "m3 train", "m3 test"))
```

### ROC curves

```{r}
my_roc_auc <- function(truth, estimate) {
    roc_auc(tibble(truth = factor(truth, levels = c(0, 1)),
                   estimate = round(estimate)),
            truth = "truth", 
            estimate,
            estimator = "binary")
}


model_roc_auc <- bind_rows(
    my_roc_auc(train_data$result, m1_train_pred),
    my_roc_auc(test_data$result, m1_test_pred),
    my_roc_auc(train_data$result, m2_train_pred),
    my_roc_auc(test_data$result, m2_test_pred),
    my_roc_auc(train_data$result, m3_train_pred),
    my_roc_auc(test_data$result, m3_test_pred)
) %>%
    mutate(name = c("m1 train", "m1 test", "m2 train", "m2 test", "m3 train", "m3 test"))
model_roc_auc
```

```{r}
my_roc_curve <- function(truth, estimate) {
    roc_curve(tibble(truth = factor(truth, levels = c(0, 1)),
                 estimate = estimate),
          estimate,
          truth = truth)
}


bind_rows(
    my_roc_curve(train_data$result, m1_train_pred) %>%
        add_column(name = "m1 train"),
    my_roc_curve(test_data$result, m1_test_pred) %>%
        add_column(name = "m1 test"),
    my_roc_curve(train_data$result, m2_train_pred) %>%
        add_column(name = "m2 train"),
    my_roc_curve(test_data$result, m2_test_pred) %>%
        add_column(name = "m2 test"),
    my_roc_curve(train_data$result, m3_train_pred) %>%
        add_column(name = "m3 train"),
    my_roc_curve(test_data$result, m3_test_pred) %>%
        add_column(name = "m3 test")
) %>%
    left_join(model_roc_auc, by = "name") %>%
    mutate(label = paste0(name, " (AUC = ", round(.estimate, 3), ")")) %>%
    ggplot(aes(x = 1 - specificity, y = sensitivity)) +
    geom_line(aes(group = name, color = label), size = 1) +
    scale_color_brewer(palette = "Set2") +
    scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0.01, 0.01))) +
    scale_y_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.01))) +
    theme(
        legend.position = c(0.75, 0.23)
    ) +
    coord_fixed() +
    labs(title = "ROC for logisitc models of volleyball statistics",
         subtitle = "Model 3 includes interaction terms for match duration.",
         color = NULL)
```

## Conclusion

Including the interactions with duration improved the model without causing overfitting.
This is not suprising as it is reasonable that the number of errors might increase in longer games and that longer games have more time for attacks, digs, etc.

Instead of the interaction terms, I could have divided the number of gameplay moves by the duration (e.g. number of attacks per minute).
However, I think the interaction term is more informative than the rate of a gameplay statistic, because, as noted above, the rate could be different depending on the duration.
If I had more time, this would be worth exploring further.
