---
title: "Volcano Eruptions"
author: "Joshua Cook"
date: "May 12, 2020"
output: github_document
---

## Setup

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")

library(glue)
library(naniar)
library(lubridate)
library(patchwork)
library(magrittr)
library(easystats)
library(MASS)
library(tidyverse)
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

theme_set(theme_minimal())
```

## Data

```{r}
volcano <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv")
eruptions <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv")
# events <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv")
# tree_rings <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/tree_rings.csv")
# sulfur <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/sulfur.csv")


volcano %<>% janitor::clean_names()
eruptions %<>% janitor::clean_names()
```

Filter only for rows with Volcano Explosivity Index (VEI) values and an eurption category that shows the eruption was confirmed.
Transform the date information into `date` objects and measure the duration of the eruptions.

```{r}
eruption_data <- eruptions %>%
    filter(!is.na(vei) & !is.na(eruption_category)) %>%
    filter(eruption_category == "Confirmed Eruption") %>%
    filter(!is.na(start_day) & start_day > 0 & start_year > 1000) %>%
    filter(!is.na(end_day) & end_day > 0) %>%
    mutate(start_date = paste(start_year, start_month, start_day, sep = "-"),
           start_date = ymd(start_date),
           end_date = paste(end_year, end_month, end_day, sep = "-"),
           end_date = ymd(end_date),
           duration = end_date - start_date) %>%
    select(volcano_number:vei, latitude:duration)
eruption_data
```

```{r}
volcano_data <- volcano %>%
    select(volcano_name, volcano_number, primary_volcano_type, country, 
           tectonic_settings, major_rock_1) %>%
    distinct()
```

```{r}
eruption_data <- inner_join(eruption_data, volcano_data, 
                by = c("volcano_number", "volcano_name"))
```

```{r}
miss_var_summary(eruption_data)
```

## EDA

```{r}
eruption_counts <- eruption_data %>%
    count(volcano_name, sort = TRUE)

eruption_data %>%
    filter(volcano_name %in% eruption_counts$volcano_name[1:10]) %>%
    ggplot(aes(x = vei, y = volcano_name)) +
    geom_jitter(height = 0.2, alpha = 0.5) +
    theme(
        legend.position = "none"
    ) +
    labs(x = "volcano explosivity index",
         y = NULL,
         title = "Explosivity of the top-10 most frequently erupting volcanoes")
```

```{r}
eruption_data %>%
    ggplot(aes(x = log(as.numeric(duration) + 1))) +
    facet_wrap(~ vei) +
    geom_density() +
    labs(x = "number of days (log + 1)",
         y = "density",
         title = "Duration of eruptions, separated by VEI")
```


## Modeling

Below, I made a few linear models and played around with the functions provided by the 'easystats' suite of packages.
I didn't do any focussed analysis, I just wanted to see what was available in these pacakges.

```{r}
scale_nums <- function(x) {
    (x - mean(x)) / sd(x)
}

d <- eruption_data %>%
    mutate(days_dur = as.numeric(duration)) %>%
    select(volcano_name, volcano_number, eruption_number,
           vei, latitude, longitude, days_dur,
           primary_volcano_type, tectonic_settings, major_rock_1) %>%
    distinct() %>%
    mutate(log_days = log(days_dur + 1),
           vei = scale_nums(vei),
           latitude = scale_nums(latitude),
           longitude = scale_nums(longitude))
```


```{r}
m1 <- lm(log_days ~ vei, data = d)

model_parameters(m1)
model_performance(m1)
check_model(m1)
```

```{r}
m2 <- lm(log_days ~ vei + tectonic_settings + major_rock_1, 
         data = d)

model_parameters(m2)
model_performance(m2)
check_model(m2)
```

```{r}
m3 <- lm(log_days ~ vei + latitude*longitude, data = d)

model_parameters(m3)
model_performance(m3)
check_model(m3)
```

```{r}
m4 <- lm(log_days ~ vei + latitude + longitude, data = d)

model_parameters(m4)
model_performance(m4)
check_model(m4)
```

```{r}
m5 <- lm(log_days ~ vei + tectonic_settings + major_rock_1 + latitude*longitude, 
         data = d)

model_parameters(m5)
model_performance(m5)
check_model(m5)
```

```{r}
compare_performance(m1, m2, m3, m4, m5)
```


```{r}
plot(compare_performance(m1, m2, m3, m4, m5))
```

From the 'insight' package.

```{r}
get_parameters(m5)
find_algorithm(m5)
find_formula(m5)
```

From the 'parameters' package.

```{r}
select_parameters(m5) %>%
    model_parameters()
```

Try to improve model 5 using the `stepAIC()` algorithm from the 'MASS' package.

```{r}
m6 <- stepAIC(m5)
```

```{r}
compare_performance(m5, m6)
plot(compare_performance(m5, m6))
```

## Mixed effect model

```{r}
library(lme4)
```
```{r}
me1 <- lmer(log_days ~ 1 + vei + (1|volcano_name), data = d)

summary(me1)
check_model(me1)
```

```{r}
me2 <- lmer(log_days ~ vei + (1|tectonic_settings/volcano_name), data = d)

summary(me2)
check_model(me2)
```

```{r}
me3 <- lmer(log_days ~ vei + (vei|volcano_name), data = d)

summary(me3)
check_model(me3)
```

```{r}
me4 <- lmer(log_days ~ vei + (1|tectonic_settings), data = d)

summary(me4)
check_model(me4)
```

```{r}
compare_performance(me1, me2, me3, me4)
```

```{r}
anova(me1, me2, me3, me4)
```

```{r}
volcanos_to_plot <- d %>%
    count(volcano_name) %>%
    top_n(9, wt = n) %>%
    pull(volcano_name)

d %>%
    mutate(pred = predict(me3)) %>%
    filter(volcano_name %in% volcanos_to_plot) %>%
    ggplot(aes(x = vei, color = volcano_name)) +
    geom_jitter(aes(y = log_days), width = 0.2, alpha = 0.5, size = 0.7) +
    geom_line(aes(y = pred), size = 1.2) +
    scale_color_brewer(palette = "Set1") +
    labs(x = "volcanic explosivity index",
         y = "days of eruption (log + 1)",
         title = "Mixed effects model of eruption duration regressed on volcanic explosivity",
         subtitle = "Each volcano was a random effect with a separate intercept and slope.")
```
```{r}
summary(me3)
```

```{r}
head(coefficients(me3)$volcano_name)
```

```{r}
ggeffects::ggpredict(me3, terms = "vei") %>%
    ggplot(aes(x = x , y = predicted)) +
    geom_ribbon(aes(ymin = predicted - std.error, ymax = predicted + std.error),
                alpha = 0.2) +
    geom_line(size = 1.3, lty = 2, color = "skyblue4") +
    geom_jitter(data = d, aes(x = vei, y = log_days, color = volcano_name), 
                size = 0.3, alpha = 0.4) +
    theme(legend.position = "none") +
    labs(x = "volcanic explosivity index",
         y = "days of eruption (log + 1)",
         title = "Mixed effects model predictions",
         subtitle = "Each volcano was a random effect with a separate intercept and slope.")
```

```{r}
stargazer::stargazer(me3, type = "text", digits = 3, digit.separator = "")
```

