---
title: "Spotify Songs"
author: "Joshua Cook"
date: "1/21/2020"
output: github_document
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")

library(conflicted)
library(patchwork)
library(GGally)
library(factoextra)
library(tictoc)
library(lubridate)
library(vip)
library(magrittr)
library(tidyverse)

conflict_prefer("filter", "dplyr")

theme_set(theme_minimal())

set.seed(0)
```

## Data

```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv') %>%
    mutate(track_album_release_date = ymd(track_album_release_date)) 
```

A few dates failed to parse because they just have the year, not month-day-year.
I do not plan to use this feature as an input for the model, so I will not worry about the parsing failures.

Some `track_ids` appear multiple times.
This is likely caused by the same song being in multiple playlists.
Annoyingly, over 1,600 songs are in multiple playlists with different genre assignments.

```{r}
# Removing playlist information removes all duplicates.
spotify_songs %>%
    select(-c(playlist_name, playlist_id, 
              playlist_genre, playlist_subgenre)) %>%
    distinct() %>%
    count(track_id) %>%
    filter(n > 1)
```


```{r}
# Removing all playlist info except for genre does not remove all duplicates.
spotify_songs %>%
    select(-c(playlist_name, playlist_id, playlist_subgenre)) %>%
    distinct() %>%
    count(track_id) %>%
    filter(n > 1)
```


## EDA

### Looking at features directly

```{r}
spotify_songs %>%
    ggplot(aes(x = track_popularity)) +
    geom_histogram(binwidth = 5,  alpha = 0.5,
                   color = "darkcyan", fill = "darkcyan") +
    scale_y_continuous(expand = expansion(mult = c(0, 0))) +
    labs(x = "track popularity",
         y = "count",
         title = "Distribution of song popularity")
```

```{r, warning=FALSE, message=FALSE}
spotify_songs %>%
    sample_n(200) %>%
    select(danceability, energy, speechiness, acousticness, 
           instrumentalness, liveness, valence) %>%
    ggpairs() +
    theme_minimal(base_size = 6)
```

```{r}
spotify_songs %>%
    group_by(playlist_genre) %>%
    summarise(n = n_distinct(track_id)) %>%
    ungroup() %>%
    ggplot(aes(x = playlist_genre, y = n)) +
    geom_col(aes(fill = playlist_genre)) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    theme(legend.position = "none") +
    labs(x = "Genre", y = "count", title = "Number of songs per genre")
```

```{r}
spotify_songs %>%
    group_by(playlist_genre, playlist_subgenre) %>%
    summarise(n = n_distinct(track_id)) %>%
    ungroup() %>%
    ggplot(aes(x = n, y = playlist_subgenre)) +
    facet_wrap(~ playlist_genre, scales = "free") +
    geom_col(aes(fill = playlist_genre)) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
    theme(legend.position = "none") +
    labs(x = "count", y = "Subgrene", title = "Number of songs per subgenre")
```

```{r}
spotify_songs %>%
    filter(!is.na(track_album_release_date)) %>%
    ggplot(aes(x = track_album_release_date, y = track_popularity)) +
    geom_point(size = 0.5, alpha = 0.2, color = "dodgerblue") +
    geom_density2d(color = "black", lty = 2, size = 1)
```

```{r}
spotify_songs %>%
    filter(!is.na(track_album_release_date)) %>%
    mutate(track_release_day = yday(track_album_release_date)) %>%
    ggplot(aes(x = track_release_day, y = track_popularity)) +
    geom_point(size = 0.5, alpha = 0.2, color = "dodgerblue") +
    geom_density2d(color = "black", lty = 2, size = 1)
```

### PCA & t-SNE

```{r}
spotify_songs %>%
    select(track_id, danceability:tempo) %>%
    skimr::skim()

spotify_songs_data <- spotify_songs %>%
    select(track_id, danceability:tempo) %>%
    distinct() %>%
    as.data.frame() %>%
    column_to_rownames("track_id")

song_pca <- prcomp(spotify_songs_data, scale = TRUE, center = TRUE)

summary(song_pca)

fviz_eig(song_pca)

tibble(
    pc = c(0, seq(1, ncol(song_pca$x))),
    cum_prop = c(0, cumsum(song_pca$sdev^2 / sum(song_pca$sdev^2)))
) %>%
    mutate(label = paste0(round(cum_prop * 100), "%")) %>%
    ggplot(aes(x = pc, y = cum_prop)) +
    geom_hline(yintercept = 0, color = "grey70") +
    geom_vline(xintercept = 0, color = "grey70") +
    geom_line(linetype = 2) +
    geom_point(size = 1.2) +
    geom_text(aes(label = label), family = "Arial", 
              nudge_x = 0.2, nudge_y = -0.03) +
    labs(x = "principal component",
         y = "cumulative proportion")

fviz_pca_var(song_pca, axes = c(1, 2))
fviz_pca_var(song_pca, axes = c(1, 3))
```

```{r}
songs_in_pca_mat <- unique(song_pca$x)

songs_in_pca <- songs_in_pca_mat %>%
    as.data.frame(stringsAsFactors = FALSE) %>%
    rownames_to_column("track_id") %>%
    as_tibble() %>%
    left_join(spotify_songs %>% select(track_id:track_popularity, 
                                       playlist_genre:playlist_subgenre),
              by = "track_id") %>%
    group_by(track_id) %>%
    filter(n() == 1) %>%
    ungroup()

songs_in_pca %>%
    sample_n(3000) %>%
    ggplot(aes(x = PC1, y = PC2, color = playlist_genre)) +
    geom_point(size = 1, alpha = 0.5) +
    labs(color = "genre",
         title = "PCA of songs by their music characteristics",
         subtitle = "Colored by genre")

songs_in_pca %>%
    sample_n(3000) %>%
    ggplot(aes(x = PC1, y = PC3, color = playlist_genre)) +
    geom_point(size = 1, alpha = 0.5) +
    labs(color = "genre",
         title = "PCA of songs by their music characteristics",
         subtitle = "Colored by genre")
```

```{r}
idx <- sample(seq(1, nrow(songs_in_pca_mat)), 5e3)
songs_tsne <- Rtsne::Rtsne(songs_in_pca_mat[idx, ])
```

```{r}
songs_tsne$Y %>%
    as.data.frame() %>%
    as_tibble() %>%
    set_names(c("Z1", "Z2")) %>%
    add_column(track_id = rownames(songs_in_pca_mat)[idx]) %>%
    left_join(spotify_songs %>% select(track_id:track_popularity, 
                                       playlist_genre:playlist_subgenre),
              by = "track_id") %>%
    group_by(track_id) %>%
    filter(n() == 1) %>%
    ungroup() %>%
    ggplot(aes(x = Z1, y = Z2, color = playlist_genre)) +
    geom_point(size = 1, alpha = 0.5) +
    labs(color = "genre", title = "t-SNE of PCA of song information")
```

## Model

I would like to get some practice with the 'tidymodels' framework, so I will use it to prepare the date (using 'rsample' and 'recipes'), instantiate a nd train a random forest model (using 'parsnip'), measure the success of the model (using 'yardstick'), and tune the model's hyperparameters (using 'tune').


For practicality reasons (as my goal is to practice using the 'tidymodels' framework, not replicate Spotify's data science team), I will restrict the data two just predicting between 3 groups

```{r}
library(tidymodels)
```

### Preparing the data

```{r}
set.seed(0)

spotify_data <- spotify_songs %>%
    filter(playlist_genre %in% c("rock", "rap", "latin")) %>%
    select(track_id, playlist_genre, danceability:tempo) %>%
    distinct() %>%
    group_by(track_id) %>%
    filter(n() == 1) %>%
    ungroup() %>% 
    group_by(playlist_genre) %>%
    sample_frac(0.7) %>%
    ungroup()

spotify_data_split <- initial_split(spotify_data, 
                                    prop = 0.8,
                                    strata = playlist_genre)
spotify_data_train <- training(spotify_data_split)
spotify_data_test <- testing(spotify_data_split)
```

```{r}
set.seed(0)
metric_pal <- randomcoloR::distinctColorPalette(11)

spotify_data_train %>%
    select(-track_id) %>%
    pivot_longer(-playlist_genre, names_to = "metric", values_to = "value") %>%
    group_by(playlist_genre, metric) %>%
    summarise(avg_value = mean(value)) %>%
    group_by(metric) %>%
    mutate(avg_value = scales::rescale(avg_value, to = c(0, 1))) %>%
    ungroup() %>%
    ggplot(aes(x = avg_value, y = metric)) +
    facet_wrap(~ playlist_genre, nrow = 1) +
    geom_col(aes(fill = metric)) +
    scale_fill_manual(values = metric_pal, guide = NULL) +
    theme(
        panel.grid.major.y = element_blank()
    ) +
    labs(x = "average value", 
         y = NULL,
         title = "Values of song metrics across genres")
```


```{r}
spotify_recipe <- recipe(playlist_genre ~ ., data = spotify_data_train) %>%
    update_role(track_id, new_role = "ID") %>%
    step_corr(all_predictors())

spotify_prep <- prep(spotify_recipe)
spotify_juiced <- juice(spotify_prep)

spotify_juiced
```

```{r}
spotify_tune_spec <- rand_forest(
    mtry = tune(),
    trees = 500,
    min_n = tune()
) %>%
    set_mode("classification") %>%
    set_engine("ranger")
```

```{r}
spotify_workflow <- workflow() %>%
    add_recipe(spotify_recipe) %>%
    add_model(spotify_tune_spec)
```



```{r}
set.seed(0)
spotify_training_folds <- vfold_cv(spotify_data_train)
spotify_training_folds
```

```{r}
doParallel::registerDoParallel()

set.seed(0)
spotify_tune_1 <- tune_grid(
    spotify_workflow,
    resamples = spotify_training_folds,
    grid = 20
)

spotify_tune_1
```

```{r}
p1 <- spotify_tune_1 %>%
    collect_metrics() %>%
    filter(.metric == "roc_auc") %>%
    ggplot(aes(x = mtry, y = min_n)) +
    geom_point(aes(color = mean, size = mean)) +
    scale_color_gradient2(low = "blue", high = "tomato", midpoint = 0.9) +
    scale_size_continuous(guide = NULL) +
    labs(title = "ROC AUC variation over the hyperparameter space")

p2 <- spotify_tune_1 %>%
    collect_metrics() %>%
    filter(.metric == "roc_auc") %>%
    select(mtry, min_n, mean) %>%
    pivot_longer(-mean, names_to = "parameter", values_to = "value") %>%
    ggplot(aes(x = value, y = mean)) +
    facet_wrap(~ parameter, nrow = 1, scales = "free") +
    geom_point(aes(color = parameter), size = 2) +
    geom_smooth(aes(group = parameter), 
                method = "loess", formula = "y ~ x", 
                color = "grey30", lty = 2, se = FALSE) +
    scale_color_brewer(type = "qual", palette = "Set2", guide = NULL)

 p1 / p2
```

```{r}
spotify_tune_grid <- grid_regular(
    mtry(range = c(2, 6)),
    min_n(range = c(5, 15)),
    levels = 5
)

spotify_tune_grid
```

```{r}
set.seed(0)

spotify_tune_2 <- tune_grid(
    spotify_workflow,
    resamples = spotify_training_folds,
    grid = spotify_tune_grid
)

spotify_tune_2
```

```{r}
spotify_tune_2 %>%
    collect_metrics() %>%
    filter(.metric == "roc_auc") %>%
    mutate(min_n = factor(min_n)) %>%
    ggplot(aes(mtry, mean, color = min_n)) +
    geom_line(alpha = 0.5, size = 1.5) +
    geom_point() +
    labs(y = "AUC")
```

```{r}
best_auc <- select_best(spotify_tune_2, "roc_auc")
spotify_rf <- finalize_model(spotify_tune_spec, best_auc)
spotify_rf
```

```{r}
spotify_rf %>%
    set_engine("ranger", importance = "permutation") %>%
    fit(playlist_genre ~ ., 
        data = select(spotify_juiced, -track_id)) %>%
    vip(geom = "point")
```

```{r}
spotify_workflow_final <- workflow() %>%
    add_recipe(spotify_recipe) %>%
    add_model(spotify_rf)

spotify_random_forest <- last_fit(spotify_workflow_final, spotify_data_split)
```


```{r}
spotify_random_forest %>%
    collect_metrics()
```

```{r}
spotify_random_forest %>%
    collect_predictions() %>%
    roc_curve(playlist_genre, .pred_latin:.pred_rock) %>%
    autoplot()
```

